--- 
title: "Machine Learning Guidelines for Natural Resource Management Practitioners"
author: "Shih-Ni Prim and Natalie Nelson"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: nataliegnelson/ml_primer
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Motivation  

As machine learning (ML) is becoming ever more powerful, some has noted that ML has not been widely used in environmental studies. Process based models might be preferred, but they can be costly and ML provides a wide variety of tools that should be explored. This booklet is meant to provide a concise guide for natural resource management practitioners. This book serves as a staring point rather than a comprehensive resource, so that practitioners can have a basic understanding of how ML works and how to utilize it to analyze data and answer research questions. We also provide case studies, R code, and links to online resources to help the readers on the journey of gaining one powerful tool that is becoming omnipresent in the research world.  


```{r eval=FALSE, include = FALSE}
# This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.
# The **bookdown** package can be installed from CRAN or Github:
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```



```{r include=FALSE}
# Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

# To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
